% <--- percent sign starts a comment line in LaTeX

%----------------------------------------------------------
% This is a sample assignment .tex file. Put your name,
% assignment number and the due date below, as shown.
% Before you typeset your own assignment try to preview 
% and print this one as follows:
%   1. Save this in a file, say hw.tex
%   2. do "latex hw"
%	3. latex will produce a pdf file that should be named hw.pdf.
%	4. Use any pdf viewer to view the document 
% ------------------------------------------------------------

\documentclass[11pt]{article}

\usepackage{fullpage,graphicx,latexsym,picinpar,amsbsy,amsmath,amsfonts}

\input{macros.tex}

\begin{document}
	
% v -- YOUR NAME and SID in the braces
\student{ Nathan Ha }{ 862377326 }  
% v -- YOUR NAME and SID in the braces
% \student{ name 2 }{sid 2 } 
% v -- ASSIGNMENT NUMBER in the braces
\assignment{ 4 } 
% v-- DUE DATE in the braces
\duedate{February 19, 2024 }  

\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\lineacross

%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\newcommand{\calT}{{\mathcal{T}}}

% PROBLEM 1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}
Give an asymptotic estimate, using the $\Theta$-notation, of the number of letters printed by the
algorithms given below. Give a complete justification for your answer, by providing an appropriate recurrence
equation and its solution.

\medskip
\noindent
(a) 
\hspace{0.01in}
%
\begin{minipage}[t]{2.4in}
\strut\vspace*{- 2.5 \baselineskip}\newline 
\input{pseudocode.tex}
\begin{program}
algorithm |PrintAs|$(n)$
   if $n\le 1$ then
      |print("A")|
   else
      for $j\assign 1$ to $n^3$
         do |print("A")|
      for $i\assign 1$ to $5$ do
         |PrintAs|$(\,\floor{n/2}\,)$
\end{program}
\end{minipage}
%
\hspace{0.4in}
(b) 
\hspace{0.01in}
%
\begin{minipage}[t]{2.4in}
\strut\vspace*{- 2.5 \baselineskip}\newline 
\input{pseudocode.tex}
\begin{program}
algorithm |PrintBs|$(n)$
   if $n\ge 4$ then
      for $j\assign 1$ to $n^2$
         do |print("B")|
      for $i\assign 1$ to $6$ do
         |PrintBs|$(\,\floor{n/4}\,)$
      for $i\assign 1$ to $10$ do
         |PrintBs|$(\,\ceiling{n/4}\,)$
\end{program}
\end{minipage}

\medskip
\noindent
(c) 
\hspace{0.01in}
%
\begin{minipage}[t]{2.4in}
\strut\vspace*{- 2.5 \baselineskip}\newline 
\input{pseudocode.tex}
\begin{program}
algorithm |PrintCs|$(n)$
   if $n\le 2$ then
      |print("C")|
   else
      for $j\assign 1$ to $n$
         do |print("C")|
      |PrintCs|$(\,\floor{n/3}\,)$
      |PrintCs|$(\,\floor{n/3}\,)$
      |PrintCs|$(\,\floor{n/3}\,)$
      |PrintCs|$(\,\floor{n/3}\,)$
\end{program}
\end{minipage}
%
\hspace{0.4in}
(d) 
\hspace{0.01in}
%
\begin{minipage}[t]{2.4in}
\strut\vspace*{- 2.5 \baselineskip}\newline 
\input{pseudocode.tex}
\begin{program}
algorithm |PrintDs|$(n)$  
   if $n\ge 5$ then
      |print("D")|
      |print("D")|
     if $(x \equiv 0 \pmod 2)$ then 
         |PrintDs|$(\,\floor{n/5}\,)$
         |PrintDs|$(\,\ceiling{n/5}\,)$
         $x\assign \ x + 3$
      else
         |PrintDs|$(\,\ceiling{n/5}\,)$
         |PrintDs|$(\,\floor{n/5}\,)$
         $x\assign 5x + 3$
\end{program}
\end{minipage}

\noindent
In part~(d), variable $x$ is a global variable initialized to $1$.
\end{problem}

%---------------------------
% solution 1
\text{}\\
The solutions for all the problems will use the master theorem, so I will just state it here. Let $a \geq 1, b > 1, c > 0$ and $d \geq 0$. If $T(n)$ satisfies the recurrence $T(n) = aT(n/b) + cn^d$.
% T(n) = {...
\[ T(x) = \begin{cases} 
   \Theta(n^{log_b{(a)}}) & a > b^d \\
   \Theta(n^dlog(n)) & a = b^d \\
   \Theta(n^d) & a < b^d 
\end{cases}
\]

\newpage

\begin{solution}
   \\
	(a) This algorithm can be rewritten as the recurrence: $A(n) = n^3 + 5A(\floor{n/2})$. Applying the master theorem, with $a = 5, b=2, c=1, d=3$. Since $a < b^d$, $T(n) = \Theta(n^3)$.
   \\\\
   (b) This algorithm can be rewritten as the recurrence: $B(n) = n^2 + 6B(\floor{n/4}) + 10B(\lceil{n/4}\rceil) = 16B(n/4) + n^2 \implies a=16, b=4, c=1, d=2$. From the master theorem, since $a=b^d$, $B(n) = \Theta(n^2\log(n))$.
   \\\\
   (c) This algorithm can be rewritten as the recurrence: $C(n) = 4C(\floor{n/3}) + n$. Applying the master theorem, with $a=4, b=3, c=1, d=1 \implies a > b^d \implies C(n) = \Theta(n^{\log_3{4}})$.
   \\\\
   (d) This algorithm can be rewritten as the recurrence: $D(n) = 2 + 2D(n/5)$. The reason for the $2D(n/5)$ is because there even though there are four recursive calls in the code, only two are executed in any given iteration. Applying the master theorem, with $a=2,b=5,c=2,d=0 \implies a > b^d$, we have $B(n) = \Theta(n^{log_5(2)})$.
\end{solution}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% PROBLEM 2
\newpage
\vspace{0.1in}
\begin{problem}
We have three sets $A$, $B$, $C$ with the following properties:

\begin{description}

\item{(a)}  $|B| = 2|A|$, $|C| = 3|A|$, 


\item{(b)} $|A\cap B| = 18$,
        $|A\cap C| = 20$,
        $|B\cap C| = 24$,

\item{(c)}
$|A\cap B\cap C| = 11$,

\item{(d)}
$|A\cup B\cup C| = 129$.

\end{description}

\noindent Use the inclusion-exclusion principle to determine the number of elements in $A$. 
Show your work.
\end{problem}


%-----------------------------

\begin{solution}
	According to the inclusion exclusion principle, we have: \\ $|A \cup B \cup C| = |A| + |B| + |C| - |A \cap B| - |A \cap C| - |B \cap C| + |A \cap B \cap C|$.
   \\
   We can substitute the values from (b), (c), (d) into the principle:
   \\
   $129 = |A| + |B| + |C| - 18 - 20 - 24 + 11 \implies |A| + |B| + |C| = 180$. To solve this, we want to put everything in terms of a single variable. From (a), we know the relationships between the three variables. I will choose to express everything in terms of $|A|$. Plugging this in, we get 
   \\ 
   $|A| +  2|A| + 3|A| = 180 \implies |A| = 30$. 

\end{solution}

% PROBLEM 3
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}
	A company, Nice Inc., will award 45 fellowships to high-achieving UCR students from four different majors: computer science, biology, political science and history. They decided to give fellowship awards to at least 8 students majoring in computer science and at most 8 biology majors. The number of political science and history majors should be between 5 and 12 students each. How many possible lists of awardees are there? You need to give a complete derivation for the final answer, using the method developed in class. 
	(Brute force listing of all lists will not be accepted.)
	
\end{problem}
	

%----------------------------
% solution 3
\begin{solution}
   Let's say that the number of CS awards is C, biology is B, political science is P, history is H. We can express the conditions as:
   \\
   $
      C \geq 8 \\
      B \leq 8 \\
      5 \leq P, H, \leq 12 \\
      C + B + P + H = 45
   $
   \\\\
   We can substitute $B' = B - 5$ and $P' = P - 5$  $(0 \leq P', H' \leq 7)$ into the sum. Also, since C has no restrictions on the upper bound, we can kind of exclude it for now\footnote[1]{As you may notice, later on when we finding the number of integer partitions, we choose k=4, which includes C. So how could we account for C in k, yet exclude it in m? This is because we are essentially setting aside 8 slots for C (which is why we subtract 8 from 45), and then seeing how many integer partitions exist. Due to the way the partition formula works, there's no reason that $B + P' + H'$ cannot actually add up to less than 27. This is because we are assuming there is a fourth group (C) that will fill that void.} (and subtract 8). We can now write:
   \\
   $B + P' + H' = 27$.
   \\\\
   We can also let the number of partitions that fit these conditions be the total number of partitions $(S())$ minus the number of invalid partitions:
   \\
   $
      S(B \leq 8 \wedge P'\leq 7 \wedge H' \leq 7) 
      \\= S() - S(B \geq 9 \lor P' \geq 8 \lor H' \geq 8) 
   $.
   \\\\
   We know that the total number of partitions, 
   \\
   S() = ${m + k - 1 \choose k - 1}, m=27, k=4 \implies S() = {30 \choose 3} = 4060$.
   \\
   m is the total number of awards, and k is the number of groups.
   \\\\
   We can also calculate the number of invalid partitions using the inclusion-exclusion principle. Also, the number of partitions with conditions is ${m - A + k - 1 \choose k - 1}$, where A is the sum of the lower bounds.
   \\
   $
      S(B \geq 9 \lor P' \geq 8 \lor H' \geq 8) = 
      \\S(B \geq 9) + S(P' \geq 8) + S(H' \geq 8) \\
      - S(B \geq 9 \land P' \geq 8)
      - S(B \geq 9 \land P' \geq 8)
      - S(B \geq 8 \land P' \geq 8)\\
      + S(B \geq 9 \land P' \geq 8 \land H' \geq 8)
      \\\\={27 - 9 + 4 - 1 \choose 4 - 1} + {27 - 8 + 4 - 1 \choose 4 - 1} + {27 - 8 + 4 - 1 \choose 4 - 1}
      \\-{27 - (9+8) + 4 - 1 \choose 4 - 1} -{27 - (9+8) + 4 - 1 \choose 4 - 1} - {27 - (8+8) + 4 - 1 \choose 4 - 1} - {27 - (9+8+8) + 4 - 1 \choose 4 - 1} 
      \\ + {27 -(8+8+9) + 4-1 \choose 4-1}
      \\=1330 + 1540 + 1540 - 286 - 286 - 364 + 10 = 3484
   $.
   \\
   Subtracting the total with the number of invalid partitions, we get $4060 - 3484 = 576$.




\end{solution}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\paragraph{Academic integrity declaration.}
I did this homework by myself. I got help from office hours, though.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}
