\documentclass{article}

\usepackage{fullpage,latexsym,picinpar,amsmath,amsfonts,graphicx}

\input{macros.tex}

\begin{document}

\centerline{\large \bf CS 111 ASSIGNMENT 3 }


\vskip 0.2in


\vskip 0.1in





%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}
\noindent a) Consider the following linear homogeneous recurrence relation: $R_n = 4R_{n-1} - 3R_{n-2}$. It is known that: $R_0 = 1$, $R_2 = 5$. Find $R_3$.

\vspace {0.1in}
\noindent b) Determine the general solution of the recurrence  equation if its characteristic equation has the following roots:  1, -2, -2, 2, 7, 7.

\vspace {0.1in}
\noindent  c) Determine the general solution of the recurrence  equation $A_n = 256A_{n-4}$.

\vspace {0.1in}
\noindent  d) Find the general form of the particular solution of the recurrence $B_n = 3B_{n-2} - 2B_{n-3}$ + 2.

\end{problem}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{problem}
Solve the following recurrence equations:

%
$a)$
\begin{eqnarray*}
        f_n &=& f_{n-1} + 4f_{n-2} + 2f_{n-3}\\
        f_0 &=& 0 \\
        f_1 &=& 1 \\
		f_2 &=& 4 
\end{eqnarray*}
%
Show your work (all steps: the characteristic polynomial and its roots, the general solution, 
using the initial conditions to compute the final solution.)


$b)$
\begin{eqnarray*}
        t_n &=& t_{n-1} + 2t_{n-2} + 2^n\\
        t_0 &=& 0 \\
        t_1 &=& 2
\end{eqnarray*}
%
Show your work (all steps: the associated homogeneous equation,
the characteristic polynomial and its
roots, the general solution of the homogeneous
equation, computing a particular solution,
the general solution of the non-homogeneous equation,
using the initial conditions to compute the final solution.)
%
\end{problem}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vskip 0.1in

\begin{problem}
We want to tile an $n\times 1$ strip with $1\times 1$ tiles that are green (G), blue (B), and red (R), $2\times 1$ purple (P) and $2\times 1$ orange (O) tiles. Green, blue and purple tiles cannot be next to each other, and there should be no two purple or three blue or green tiles in a row (for ex., GGOBR is allowed, but GGGOBR, GROPP and PBOBR are not). Give a formula for the number of such tilings. Your solution must include a recurrence equation (with initial conditions!), and a full justification. You do not need to solve it. 


\end{problem}
%%%%%%%%%%%%%%%%%%%%


\paragraph{Academic integrity declaration.}
The homework papers must include at the end an academic integrity declaration. This should be a brief paragraph where you state
\emph{in your own words}  (1) whether you did the homework individually or in collaboration with a partner student (if so, provide the name), 
and (2) whether you used any external help or resources. 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\vskip 0.1in

\paragraph{Submission.}
To submit the homework, you need to upload the pdf file to Gradescope. If you submit with a partner, you need
to put two names on the assignment and submit it as a group assignment.
\end{document}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%